Network Configuration Synthesis with Abstract Topologies

### 2016-11-14 paper/abstraction/paper.tex
update section 6 to talk about merging abstract paths

%
To infer bounds on the number of edge-disjoint paths between concrete nodes, we use information about the structure of the underlying concrete topologies encoded in abstraction constraints.
%
For each node in the abstract PG,
the idea is to maintain facts learned about the number of disjoint paths
to some number of concrete routers in the node's role.
More specifically, we maintain invariants of the form:
%
\[ \begin{array}{c}
  L_{X_1}, \ldots, L_{X_n}(j,k)
\end{array} \]
\noindent
%
where each label $L \in \{S,A\}$ is either $S$, which stands for ``some" or is $A$, which stands for ``all". There is one label for each pod in the abstraction pod hierarchy. For a given node, $L_{X_1}$ corresponds to the outermost pod, $L_{X_{n-1}}$ corresponds to the innermost pod, and $L_{X_n}$ to the the node itself.
Semantically, $L_{X_1}, \ldots, L_{X_n}(j,k)$ means that starting from some concrete source node, for some/all pods $X_1, \ldots, X_{n-1}$
and for some/all groups of nodes in the role $X_n$ of size $j$,
%, $L_{X_2}, \ldots, L_{X_n}(j,k)$ holds. The base case $L_{X_n}(j,k)$ means that for "some"/``all" groups of size $j$ concrete routers,
there are $k$ paths to each such that all $j*k$ paths are edge-disjoint.

For example,
an inference of the form $A_{S}(2,3)$ states that for \emph{all} groups of 2 nodes
in the $S$ (Spine) role, there are 6 disjoint paths to the group---3
for each of the 2 nodes. A fact such as $S_{S}(2,3)$, states that
\emph{some} group of 2 nodes has the property as opposed to all.

@@

To learn information about disjoint paths, we develop a collection of sound inference rules for invariants of this form. Figure~\ref{fig:inference-rules} shows each of the inference rules. The rules are read from bottom to top. The label on the bottom left represents a known fact of the form described above. We use $L$ to represent a rule that is parametric over the label ($S$ or $A$). The labels on other nodes correspond to facts learned after applying the inference rule. Conditions that must be \emph{valid} for the rule to apply are shown in a box next to the rule.

@@

The first rule, \infrule{I-out1} applies to a learned fact of the form $L_m(j,k)$ where the number of outgoing edges from any concrete node in the $m$ role is greater than 0. In the worst case, the largest group of concrete nodes we could hope to reach at the $n$ role would be $e_1$ since all j nodes at the bottom may have outgoing edges to the same concrete nodes at the top. Furthermore, the total number of disjoint paths to the $j$ nodes at the bottom is equal to $j*k$. Since adding edges between the nodes keeps the existing paths disjoint, and since we can not exceed the current number of disjoint paths to the concrete nodes in role $m$ on the bottom, the largest reachable group for the top node will be $min(j*k,e_1)$. Since we do not know for certain that all $j$ nodes have outgoing edges to the same nodes at the top, we conservatively use $1$ for the number of disjoint paths to each node on top.

@@

The rule \infrule{I-striping} is the most complicated case.
It starts with the fact L$_m(j,k)$ at role $m$ and can be applied if each edge multiplicity $e_i > 0$ is valid given the constraints. The first inference for role $n$ tries to find the largest reachable group with disjoint paths to each. The idea is similar to the rule \infrule{I-out1}, but is able to use the fact that $e_2 > 0$ to learn more about the structure of the concrete topology. In particular it uses the following inequality, where $g$ represents the size of the group for the role $n$:
%
\[ \begin{array}{c}
  (m-j)*e_1 \geq (n-g)*e_2
\end{array} \]
\noindent
%
The remaining nodes ($m-j$) that are not part of the reachable group in the bottom role, each have $e_1$ outgoing edges and must be able to at least "fill" the incoming edges for the remaining nodes not in the reachable group at the top ($n-g$) with $e_2$ incoming edges each. Solving the inequality gives the lower bound for $g$ used in Figure~\ref{fig:inference-rules}.

@@

Finally, the \infrule{I-local} and \infrule{I-global} rules allow us to lift each of the other rules to reason across pod hierarchies. The first rule, \infrule{I-local}, says that if you have an inference from L$_m(j,k)$ to L$_n(j',k')$ on the left, then if both nodes are in the same local pod P, you can leave all labels up to P unchanged. The global rule \infrule{I-global} says the same thing except when the edge goes between pods, you can infer the fact for all pods Q since the multiplicities apply equally to each pod.

@@

The abstract disjoint path analysis starts from a fixed source location $src$ and repeatedly applies the inference rules from Figure~\ref{fig:inference-rules} until it reaches a fixed point. Inference rule conditions are checked for validity over all concrete topologies using an SMT solver. In order to ensure termination, we make the observation that, for any fact learned of the form $L(j,k)$, it is sound to instead infer $L(j',k')$ if $j' \leq j$ and $k' \leq k$. Therefore, for each inference rule that learns a fact of the form $L(j,k)$ we minimize the symbolic expressions for $j$ and $k$ subject to the topology constraints using the optimizing constraint solver $\nu$Z~\cite{z3opt}.

>>>>

%
To infer bounds on the number of edge-disjoint paths between concrete nodes, we use information about the structure of the underlying concrete topologies encoded in abstraction constraints.
%
For each node in the abstract PG,
the idea is to maintain facts learned about the number of disjoint paths
to groups of concrete routers in the node's role.
More specifically, we maintain invariants of the form:
%
\[ \begin{array}{c}
  L_{X_1}, \ldots, L_{X_n}(j,k)
\end{array} \]
\noindent
%
where each label $L \in \{S,A\}$ is either $S$, which stands for ``some" or is $A$, which stands for ``all". There is one label for each pod in the abstraction pod hierarchy under which the abstract node appears. For a given node, $L_{X_1}$ corresponds to the outermost pod, $L_{X_{n-1}}$ corresponds to the innermost pod, and $L_{X_n}$ to the the node itself.
Semantically, $L_{X_1}, \ldots, L_{X_n}(j,k)$ means that starting from some concrete source node, for some/all pods $X_1, \ldots, X_{n-1}$
and for some/all groups of nodes in the role $X_n$ of size $j$,
%, $L_{X_2}, \ldots, L_{X_n}(j,k)$ holds. The base case $L_{X_n}(j,k)$ means that for "some"/``all" groups of size $j$ concrete routers,
there are $k$ paths to each such that all $j*k$ paths are edge-disjoint.

For example,
an inference of the form $A_{Q} A_{TL}(2,3)$ states that, from some fixed source location, for \emph{all} pods Q, and \emph{all} groups of 2 nodes in the TL role, there are 6 disjoint paths to the group---3 for each of the 2 nodes.

@@

To learn information about disjoint paths, we develop a collection of sound inference rules for invariants of the above form. Figure~\ref{fig:inference-rules} shows each of the inference rules. The rules are read from bottom to top. The label on the bottom left represents a known invariant. We use $L$ to represent a rule that is parametric over the label ($S$ or $A$). The labels on other nodes correspond to new invariants learned after applying the inference rule. Conditions that must be \emph{valid} for the rule to apply are shown in a box next to the rule.

@@

The first rule, \infrule{I-out1} applies to a learned invariant of the form $L_m(j,k)$ where the number of outgoing edges from any concrete node in the $m$ role is greater than 0. In the worst case, the largest group of concrete nodes we could hope to reach at the $n$ role would be $e_1$ since all j nodes at the bottom may have outgoing edges to the same concrete nodes at the top. Furthermore, the total number of disjoint paths to the $j$ nodes at the bottom is equal to $j*k$. Since adding disjoint edges between the nodes keeps the existing paths disjoint, and since we can not exceed the current number of disjoint paths to the concrete nodes in role $m$ on the bottom, the largest reachable group for the role $n$ on the top will be $min(j*k,e_1)$. However, because we do not know for certain that all $j$ nodes in role $m$ have outgoing edges to the same nodes at the top, we conservatively use $1$ for the number of disjoint paths to each node on top.

@@

The rule \infrule{I-striping} is the most complicated case.
It starts with the invariant L$_m(j,k)$ at role $m$ and can be applied if each edge multiplicity $e_i > 0$ is valid given the constraints. The first inference for role $n$ tries to find the largest reachable group with disjoint paths to each. The idea is similar to the rule \infrule{I-out1}, but is able to use the fact that $e_2 > 0$ to learn more about the structure of the concrete topology. In particular it uses the following inequality, where $g$ represents the size of the group for the role $n$:
%
\[ \begin{array}{c}
  (m-j)*e_1 \geq (n-g)*e_2
\end{array} \]
\noindent
%
The remaining nodes ($m-j$) that are not part of the reachable group in the bottom role, each have $e_1$ outgoing edges and must be able to at least "fill" the incoming edges for the remaining nodes not in the reachable group at the top ($n-g$) with $e_2$ incoming edges each. Solving the inequality gives the lower bound for $g$ used in Figure~\ref{fig:inference-rules}.

@@

Finally, the \infrule{I-local} and \infrule{I-global} rules allow us to lift each of the other rules to reason across pod hierarchies. The first rule, \infrule{I-local}, says that if there is an inference from L$_m(j,k)$ to L$_n(j',k')$ on the left, then if both nodes are in the same local pod P, we can leave all labels up to P unchanged. The global rule \infrule{I-global} says that when the edge goes between pods, we can infer the fact for all pods Q since the multiplicities apply uniformly for each pod.

@@

The abstract disjoint path analysis starts from a fixed source location $src$ and repeatedly applies the inference rules from Figure~\ref{fig:inference-rules} until it reaches a fixed point. Inference rule conditions are checked for validity over all concrete topologies using an SMT solver. In order to ensure termination, we make the observation that, for any fact learned of the form $L(j,k)$, it is sound to instead infer $L(j',k')$ if $j' \leq j$ and $k' \leq k$. Therefore, for each inference $L(j,k)$ we minimize the symbolic expressions for $j$ and $k$ subject to the topology constraints using the optimizing SMT solver $\nu$Z~\cite{z3opt}.

The algorithm associates with each inference, the set of abstract locations used to derive the inference, \IE, an overapproximation of set of concrete locations used for the edge-disjoint paths. When applying an inference rule between an invariant at role $m$, with a set of abstract locations $C$, to learn an invariant at role $n$, we associate the new invariant with the set $C \cup \{ n \}$. 
%
When, at a single PG node there is an inference $L_{X_1} \dots L_{X_{n-1}}$ A$_m(j,k)$ with location set $C$ and inference $L_{X_1} \dots L_{X_{n-1}}$ L$_m(j',k')$ with location set $D$, where $C \cap D = \emptyset$, we add a new inference with location set $C \cup D$: 
%
$$L_{X_1} \dots L_{X_{n-1}} L_m(\min(j, j'), k + k')$$ 
%
\noindent
The idea is that, because non-overlapping abstract paths can never contain overlapping concrete paths, due to the graph homomorphism property, we can merge inferences about these disjoint paths in the abstraction.
